Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction

### 2018-01-26 paper/manuscript_revised_jan25.tex
Eliminate repition in AprioriAll description

\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{bissonnj, marofkec, coxt3\}@myumanitoba.ca}}

@@

@@

The primary baseline of this work is \cite{major-hurricane-model}, which uses a specific implementation of AprioriAll (first proposed in \cite{AprioriAll-original}) to determine frequent hurricane trajectory patterns. These trajectory patterns correspond to interesting trajectory association rules, which are used to train a prediction model. The implementation of AprioriAll used in \cite{major-hurricane-model} also serves as the foundation for the model proposed in this work. AprioriAll transforms a transaction database into a \textit{customer} database, where each entry contains a customer and an ordered sequence of purchased items. The ordered sequence database is then mined for frequent ordered patterns, or frequent sequences. The modification proposed in \cite{major-hurricane-model} directly mines frequent ordered patterns and thus skips the transformation step of AprioriAll.

@@

The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed. By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved. This includes coordinates which may ultimately be infrequent. The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences. This allows the prepared (mapped) database to be mined directly without the need to complete the mapping step every time AprioriAll is executed.

@@

Figure (1) shows an example of how the trajectory database is structured after the mapping preprocessing step. Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates. All hurricanes in HURDAT2 have an ID corresponding to the location (i.e. the Atlantic), the index ordered by appearance in the hurricane year, and the year itself. For example, hurricane Katrina of year 2005 corresponds to AL122005. The preprocessing step proposed is executed exactly once, allowing the hurricane database to be mined without the initial grouping and mapping stage.

@@

There are multiple areas to explore in the hurricane trajectory prediction problem space with respect to future work. Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched. While \cite{major-hurricane-model} proposed an extrapolation method to recover from rule-miss-incidents, more work must be completed to reduce the number of non-matchable trajectories. Many of the mismatched trajectories are outliers with very little conformance to existing hurricane trajectories as seen in figure (3). Additionally, more experiments will be run on different inputs including training data from other regions such as the Pacific region. Further approaches to increasing the correctness ratio of WARD-HTP will also be investigated. This includes the investigation of alternative methods for defining training data weight.

>>>>

\author{\IEEEauthorblockN{Taylor Cox, Caden Marofke}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{coxt3, marofkec\}@myumanitoba.ca}}

@@

% Reduce repition: highlight modificatiosn of aprioriall less frequently
% Clearly state that weighted support is our idea, inspired by some work but we have a diff approach

@@

The primary baseline of this work is \cite{major-hurricane-model}, which uses a specific implementation of AprioriAll (first proposed in \cite{AprioriAll-original}) to determine frequent hurricane trajectory patterns. These trajectory patterns correspond to interesting trajectory association rules, which are used to train a prediction model. The implementation of AprioriAll used in \cite{major-hurricane-model} also serves as the foundation for the model proposed in this work. AprioriAll transforms a transaction database into a \textit{customer} database, where each entry contains a customer and an ordered sequence of purchased items. The ordered sequence database is then mined for frequent ordered sequences.

@@

The hurricane trajectory model in this work does not implement step (1), as the training data from HURDAT2 was instead translated into a $\langle HurricaneID, CoordinateSequence\rangle$ database as a preprocessing step before any aspects of the model are executed. Figure (1) shows an example of how the trajectory database is structured after the initial mapping stage. Only two dimensions are needed in the database, being the hurricane ID and the sequence of latitude-longitude coordinates. All hurricanes in HURDAT2 have an ID corresponding to the location (i.e. the Atlantic), the index ordered by appearance in the hurricane year, and the year itself. For example, hurricane Katrina of year 2005 corresponds to AL122005. By directly translating the data instead of processing it with AprioriAll, all trajectories in the original database are preserved. This includes coordinate sequences which may ultimately be infrequent. The sequential mapping of AprioriAll is performed as a preprocessing step so that the database used in the prediction model will always consist of hurricanes and their coordinate sequences. The prepared (mapped) database may therefore be mined directly without the need for repeated mapping in subsequent executions of AprioriAll.

@@

@@

There are multiple areas to explore in the hurricane trajectory prediction problem space with respect to future work. Experiments in this work show that all datasets contain a subset of trajectories for which no association rule could be matched. While \cite{major-hurricane-model} proposed an extrapolation method to recover from rule-miss incidents, more work must be completed to reduce the number of non-matchable trajectories. Many of the mismatched trajectories are outliers with very little conformance to existing hurricane trajectories as seen in figure (3). Additionally, more experiments will be run on different inputs including training data from other regions such as the Pacific region. Further approaches to increasing the correctness ratio of WARD-HTP will also be investigated. This includes the investigation of alternative methods for defining training data weight.

### 2018-01-26 winter_term/manuscript_revised_jan25.tex
Eliminate repition in AprioriAll description

\author{\IEEEauthorblockN{Taylor Cox, Caden Marofke}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{coxt3, marofkec\}@myumanitoba.ca}}

>>>>

\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{bissonnj, marofkec, coxt3\}@myumanitoba.ca}}
